### 2020-09-02 src/uk.Rmd
all but .RData

---
title: "UK"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../outputs")
  })
---

```{r}
library(rmarkdown)
library(data.table)
library(tidyverse)
library(readxl)
library(httr)
library(rprojroot)

library(ggplot2)
#library(geojsonio)
library(rgdal)
library(maptools)
library(sp)
library(sf)
library(rgeos)
#library(classInt)
#library(RColorBrewer)
library(knitr)
#library(viridis)
#library(hrbrthemes)
library(fs)
library(betareg)
library(broom)

library(AER)
library(MASS)
#library(foreign)
library(stargazer)
library(ivpack) 
library(leaflet)
library(rvest)
library(xml2)
library(tmaptools)
library(spdep)
library(spatialreg)
library(spatialEco)

# This is the project path
path <- find_rstudio_root_file()
```

## Merge bua with buasd

```{r}
# Get BUA 
# Some BUA also include sub-division, so the two spatial objects need to be downloaded selerately and merged:

# (i) directly from the web
# buasd <- readOGR("https://opendata.arcgis.com/datasets/bc2dbefc1285410e9fdf800745d1280f_0.geojson") #, layer="OGRGeoJSON")
# bua <- readOGR("https://opendata.arcgis.com/datasets/210f705366044b8c9ca3fa6f4f83fa5d_0.geojson")

# or, (ii) from the geojson saved locally
path.json <- paste(path, "/data/uk/raw/Built-up_Area_Sub_Divisions_(December_2011)_Boundaries.geojson", sep = "")
buasd <- readOGR(path.json)#, layer="OGRGeoJSON")
path.json <- paste(path, "/data/UK/raw/Built-up_Areas_(December_2011)_Boundaries_V2.geojson", sep = "")
bua <- readOGR(path.json)#, layer="OGRGeoJSON")
# source: https://geoportal.statistics.gov.uk/

# spatial transformations
buasd <- spTransform(buasd, CRS("+init=epsg:4326"))
bua <- spTransform(bua, CRS("+init=epsg:4326"))

dim(bua) # 5830
dim(buasd) #1826

# drop the BUA with sub-division to only include the sub-divisions (see rbind() below)
bua <- bua[bua$has_sd=="N",] #5262

names(bua)
names(buasd)

bua$has_sd <- NULL
bua$urban_bua <- NULL
bua$sd_count <- NULL
buasd$buasd_id <- NULL
names(buasd)[3] <- "bua11cd"
names(buasd)[4] <- "bua11nm"
dim(bua)
dim(buasd)
bua$sd_count <- NULL
buasd$sd_type <- NULL

# merge bua with buasd
bua_buasd <- rbind(bua, buasd, makeUniqueIDs = TRUE)
dim(bua_buasd@data)

# factors to characters
bua_buasd@data$bua11cd <- as.character(bua_buasd@data$bua11cd)
bua_buasd@data$bua11nm <- as.character(bua_buasd@data$bua11nm)
```

## Broadband speed from Speedchecker Ltd., 2017

```{r}
# load broadband data
path.data <- paste(path, "/data/uk/raw/broadband.speed/2015-2017/broadbandspeed.csv", sep = "")
bb2017 <- read_csv(path.data, col_names = T)
range(bb2017$DateTimeStamp)

bb2017 <- bb2017 %>% 
  filter(DateTimeStamp > "2017-01-01") %>% 
  filter(DownloadSpeed > 512 & DownloadSpeed< 362000) %>% # Riddlesden and Singleton 2014 + my interpretation
  filter(UploadSpeed > 100 & UploadSpeed < 21000) %>% # Riddlesden and Singleton 2014 + my interpretation
  drop_na()

# convert points to spatial object
coords_bb <- cbind(bb2017$Longitude, bb2017$Latitude)
bb.bua <- SpatialPointsDataFrame(coords_bb, data = data.frame(bb2017))
proj4string(bb.bua) <- CRS("+init=epsg:4326") #define projection

# spatial join to bua_buasd
bb.bua.sp <- over(bb.bua, bua_buasd[, "bua11cd"]) # not a spatial object
bb.bua$bua11cd <- bb.bua.sp$bua11cd

# create dataframe object to analyse (remove lat and lon?)
bb2017 <- bb.bua@data

# drop NAs
sapply(bb2017, function(x) sum(is.na(x)))
bb2017 <- bb2017[!is.na(bb2017$bua11cd),]
#NA bua11cd, coord outside bua, buasd or the UK = 476381 / 1953654 obs (76% of obs retained)

bb2017 <- bb2017 %>%
  group_by(bua11cd) %>% # this also include buasd
  summarize(download = median(DownloadSpeed), upload = median(UploadSpeed), n.tests = n()) 
```

## Broadband speed from Speedchecker Ltd., 2011

```{r}
# load broadband data
path.data <- paste(path, "/data/uk/raw/broadband.speed/2011", sep = "")
files <- dir_ls(path=path.data, pattern="*.csv") # dir_ls() also inludes the paths so it works for .Rmd

bb2011 <- lapply(files, function(x) read_delim(x, delim=",", col_names = F)) %>% 
  bind_rows() #
names(bb2011)[1:6] <- c("DateTimeStamp","DownloadSpeed","UploadSpeed","provider","Latitude", "Longitude")

bb2011 <- bb2011 %>% 
  filter(DownloadSpeed > 512 & DownloadSpeed< 362000) %>% # Riddlesden and Singleton 2014 + my interpretation
  filter(UploadSpeed > 100 & UploadSpeed < 21000) %>% # Riddlesden and Singleton 2014 + my interpretation
  drop_na()

# convert points to spatial object
coords_bb <- cbind(bb2011$Longitude, bb2011$Latitude)
bb.bua <- SpatialPointsDataFrame(coords_bb, data = data.frame(bb2011))
proj4string(bb.bua) <- CRS("+init=epsg:4326") #define projection

# spatial join to bua_buasd
bb.bua.sp <- over(bb.bua, bua_buasd[, "bua11cd"]) # not a spatial object
bb.bua$bua11cd <- bb.bua.sp$bua11cd

# create dataframe object to analyse (remove lat and lon?)
bb2011 <- bb.bua@data

# drop NAs
sapply(bb2011, function(x) sum(is.na(x)))
bb2011 <- bb2011[!is.na(bb2011$bua11cd),]
#NA bua11cd, coord outside bua, buasd or the UK = 269310 / 1307293 obs (80% of obs retained)

bb2011 <- bb2011 %>%
  group_by(bua11cd) %>% # this also include buasd
  summarize(download = median(DownloadSpeed), upload = median(UploadSpeed), n.tests = n()) 
```

## Population, population rank and differences; LHS variable

```{r}
pop <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_2010_1.data.csv?geography=1119879169...1119882228,1119882230...1119885230,1119885232...1119885236,1119885238...1119885256,1119885263...1119885265,1119885267,1119885257...1119885262,1119885266,1119885268...1119885792&gender=0&c_age=200&measures=20100&signature=NPK-8f824e8ae641543e1e4ae2:0xa7a1e65a29f4c48e3059704c1fe86a40e2c0b670")
# saved also as a wide .csv in /ict_un_us_uk/data/uk/raw/bua_pop.csv

pop <- pop %>% 
  dplyr::select(DATE, GEOGRAPHY_CODE, GEOGRAPHY_NAME, OBS_VALUE) %>%
  spread(DATE, OBS_VALUE) %>% 
  mutate(r2011 = dense_rank(desc(`2011`))) %>% 
  mutate(r2018 = dense_rank(desc(`2018`))) %>%
  mutate(r_diff = r2011 - r2018) %>%
  mutate(r_diff_tr = (r_diff + sum(!is.na(r_diff)))/(2*sum(!is.na(r_diff)))) %>%
  rename(pop2011 = `2011`) %>%
  rename(pop2012 = `2012`) %>%
  rename(pop2013 = `2013`) %>%
  rename(pop2014 = `2014`) %>%
  rename(pop2015 = `2015`) %>%
  rename(pop2016 = `2016`) %>%
  rename(pop2017 = `2017`) %>%
  rename(pop2018 = `2018`)
# for normalisation N = sum(!is.na(r_diff))
# Check later if this correct for the regression data

# Scotland is excluded as the methodology for Scottish Settlements and Localities is different:
# https://www.nrscotland.gov.uk/statistics-and-data/statistics/statistics-by-theme/population/population-estimates/special-area-population-estimates/settlements-and-localities/background-information
# pop %>% filter(str_detect(GEOGRAPHY_CODE, "S")) 
```

## IMD NOT USED

```{r}
# #imd.en.2019 <- read_csv(unzip("https://data.cdrc.ac.uk/system/files/English%20IMD%202019.zip" )
# temp <- tempfile()
# download.file("https://data.cdrc.ac.uk/system/files/English%20IMD%202019.zip",temp)
# imd.en.2019 <- read.table(unz(temp, "English"))
# unlink(temp)

# England 2010
#england.imd2010 <- read_csv(url("https://assets.publishing.service.gov.uk/government/uploads/system/uploads/attachment_data/file/15240/1871702.csv"))
# also saved as imd_england2010.csv

# England 2019
#england.imd2019 <- read_csv("https://assets.publishing.service.gov.uk/government/uploads/system/uploads/attachment_data/file/845345/File_7_-_All_IoD2019_Scores__Ranks__Deciles_and_Population_Denominators_3.csv")

# Wales 2011
#url <- "https://gov.wales/sites/default/files/statistics-and-research/2019-04/wimd-2011-individual-domain-scores-and-overall-index-scores-for-each-lsoa.xls"
#GET(url, write_disk(tf <- tempfile(fileext = ".xls")))
#wales.imd2011 <- read_excel(tf, 2L)
# also saved as wimd-2011-individual-domain-scores-and-overall-index-scores-for-each-lsoa.xls

# Wales 2019
#url <- "https://gov.wales/sites/default/files/statistics-and-research/2019-12/wimd-2019-index-and-domain-scores-by-small-area_0.ods"
#GET(url, write_disk(tf <- tempfile(fileext = ".xls")))
#wales.imd2019 <- read_ods(tf, 2L, skip = 2)
# also saved as wimd-2019-index-and-domain-scores-by-small-area_0.ods

```

## Jobseeker's Allowance with rates and proportions NOT USED

```{r}
# jobseeker <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_1_1.data.csv?geography=1119879169...1119882228,1119882230...1119885230,1119885232...1119885236,1119885238...1119885256,1119885263...1119885265,1119885267,1119885257...1119885262,1119885266,1119885268...1119885792&date=latestMINUS108,latestMINUS96,latestMINUS84,latestMINUS72,latestMINUS60,latestMINUS48,latestMINUS36,latestMINUS24,latestMINUS12&sex=7&item=1&measures=20100&signature=NPK-8f824e8ae641543e1e4ae2:0xbd6481e37da64ded2c7523fd7226460d3db46f0f")
# # saved as wide .csv in \ict_un_us_uk\data\uk\raw\job_seeker_allowance.csv
# 
# jobseeker <- read_json("https://www.nomisweb.co.uk/api/v01/dataset/NM_1_1.jsonstat.json?geography=1119879169...1119882228,1119882230...1119885230,1119885232...1119885236,1119885238...1119885256,1119885263...1119885265,1119885267,1119885257...1119885262,1119885266,1119885268...1119885792&date=latestMINUS108,latestMINUS96,latestMINUS84,latestMINUS72,latestMINUS60,latestMINUS48,latestMINUS36,latestMINUS24,latestMINUS12&sex=7&item=1&measures=20100&signature=NPK-8f824e8ae641543e1e4ae2:0xea3fd585c7b9709c1f247b9a304c1355a9b5e13a")
# 
# jobseeker %>% spread_all
# 
# jobseeker_ <- jobseeker %>% 
#   dplyr::select(DATE, GEOGRAPHY_CODE, GEOGRAPHY_NAME, OBS_VALUE) %>%
#   spread(DATE, OBS_VALUE) #%>% 
#   mutate(r2011 = dense_rank(desc(`2011`)))

```

## Population density

```{r}
pop.dens <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_143_1.data.csv?date=latest&geography=1119879169...1119882228,1119882230...1119885230,1119885232...1119885236,1119885238...1119885256,1119885263...1119885265,1119885267,1119885257...1119885262,1119885266,1119885268...1119885792&rural_urban=0&cell=0,1&measures=20100&signature=NPK-8f824e8ae641543e1e4ae2:0xbd7005144df55f1693a38d07f3e255171c673600")
# saved also as pop_dens.csv

pop.dens <- pop.dens %>% 
  dplyr::select(GEOGRAPHY_CODE, GEOGRAPHY_NAME, CELL_NAME, OBS_VALUE) %>%
  spread(CELL_NAME, OBS_VALUE) %>% 
  mutate(pop.dens = `All usual residents` / `Area Hectares`) %>%
  dplyr::select(GEOGRAPHY_CODE, pop.dens)
```

## urban / rural classification

```{r}
# England and Wales 2011 urban/rural classification for OA
# http://geoportal.statistics.gov.uk/datasets/rural-urban-classification-2011-of-output-areas-in-england-and-wales
path.data <- paste(path, "/data/uk/raw/RUC11_OA11_EW.zip", sep = "")
ur.ew <- read_csv(unzip(path.data, "RUC11_OA11_EW.csv"))

# NOT USED
# Scotland
# The Scottish Government's Urban Rural Classification 2013/14, 2011 Output Area
# https://www.isdscotland.org/Products-and-Services/GPD-Support/Geography/Urban-Rural-Classification/
# classification info: https://www2.gov.scot/Topics/Statistics/About/Methodology/UrbanRuralClassification
# ur.sc <- read_csv("https://www.isdscotland.org/Products-and-Services/GPD-Support/Geography/Urban-Rural-Classification/_docs/2013_2014_urban_rural/oa2011_urban_rural_2013_2014.csv")

# lookup file for OA-to-BUA/BUASD
lookup.out.bua <- read_csv("https://opendata.arcgis.com/datasets/edaf7c4b5e6e401d987c30c4de6b63e6_0.csv")
# saved as Output_Area__2011__to_Built-up_Area_Sub-division_to_Built-up_Area_to_Local_Authority_District_to_Region__December_2011__Lookup_in_England_and_Wales.csv
dim(lookup.out.bua)
lookup.out.bua <- lookup.out.bua %>% dplyr::select(OA11CD, BUASD11CD, BUASD11NM, BUA11CD, BUA11NM)

ur.ew <- merge(ur.ew, lookup.out.bua, by = "OA11CD", all.x = T)
sapply(ur.ew, function(x) sum(is.na(x)))

# create new id which contains BUA or BUASD id
ur.ew$BUA_SD_CD <- ifelse(is.na(ur.ew$BUASD11CD), ur.ew$BUA11CD, ur.ew$BUASD11CD) #replace BUA sub divisions
ur.ew$BUA_SD_NM <- ifelse(is.na(ur.ew$BUASD11NM), ur.ew$BUA11NM, ur.ew$BUASD11NM) #replace BUA sub divisions
# with BUA for those BUA which don't have sub divisions
# sum(ur.ew$BUASD11NM==ur.ew$BUA11NM) #check

length(unique(ur.ew$BUA_SD_CD)) #6621, 6622 now 
length(unique(ur.ew$BUA_SD_NM)) #6621, 6622 now

ur.ew <- ur.ew %>% 
  filter(!is.na(BUA11CD) | !is.na(BUASD11CD) | !is.na(BUA_SD_CD)) %>%
  mutate(urban = gsub("([A-Za-z]+).*", "\\1", RUC11)) %>% # take the first word from split string
  dplyr::select(BUA_SD_CD, urban) %>%
  distinct() # this leads to 6621 unique obs., so every BUA_SD_CD has a unique value for urban
```

## Ethnicity in 2011

```{r}
british.pop <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_574_1.data.csv?date=latest&geography=1119879169...1119882228,1119882230...1119885230,1119885232...1119885236,1119885238...1119885256,1119885263...1119885265,1119885267,1119885257...1119885262,1119885266,1119885268...1119885792&rural_urban=0&cell=0,1&measures=20100&signature=NPK-8f824e8ae641543e1e4ae2:0xadba1f3dfe1e84341e0d33f6d082210022a1418e")
# saved as british_pop.csv

british.pop <- british.pop %>% 
  dplyr::select(GEOGRAPHY_CODE, GEOGRAPHY_NAME, CELL_NAME, OBS_VALUE) %>%
  spread(CELL_NAME, OBS_VALUE) %>% 
  rename(all.pop = `All categories: Ethnic group`) %>%
  rename(british.pop = `English/Welsh/Scottish/Northern Irish/British`) %>%
  mutate(british.pop.share = british.pop / all.pop) %>%
  dplyr::select(GEOGRAPHY_CODE, british.pop.share)
```

## Working from home 2011

```{r}
work.home <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_153_1.data.csv?date=latest&geography=1119879169...1119882228,1119882230...1119885230,1119885232...1119885236,1119885238...1119885256,1119885263...1119885265,1119885267,1119885257...1119885262,1119885266,1119885268...1119885792&rural_urban=0&cell=0,9&measures=20100&signature=NPK-8f824e8ae641543e1e4ae2:0xfa76d191fd681a934f45a8a0f35bad74f3fe0877")
# save as work_home.csv

work.home <- work.home %>% 
  dplyr::select(GEOGRAPHY_CODE, GEOGRAPHY_NAME, CELL_NAME, OBS_VALUE) %>%
  spread(CELL_NAME, OBS_VALUE) %>% 
  rename(working.pop = `All categories: Distance travelled to work`) %>%
  rename(work.home = `Work mainly at or from home`) %>%
  mutate(work.home.share = work.home / working.pop) %>%
  dplyr::select(GEOGRAPHY_CODE, work.home.share)
```

## Services 2011

```{r}
services <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_1020_1.data.csv?date=latest&geography=1119879169...1119882228,1119882230...1119885230,1119885232...1119885236,1119885238...1119885256,1119885263...1119885265,1119885267,1119885257...1119885262,1119885266,1119885268...1119885792&c_indgpuk11=0...8&c_sex=0&measures=20100&signature=NPK-8f824e8ae641543e1e4ae2:0xe16647b1bfd6ad3e2bbea89fa2de3e97bbbab3a9")
# saved as industries.csv

services <- services %>%
  dplyr::select(GEOGRAPHY_CODE, GEOGRAPHY_NAME, C_INDGPUK11_NAME, OBS_VALUE) %>% # C_INDGPUK11: this is the industry code
  spread(C_INDGPUK11_NAME, OBS_VALUE) %>% 
  mutate(services.share = (`G, I Distribution, hotels and restaurants` +
                             `H, J Transport and communication` +
                             `K, L, M, N Financial, Real Estate, Professional and Administrative activities` +
                             `O, P, Q Public administration, education and health` +
                             `R, S, T, U Other`) / `All categories: Industry`)  %>%
  dplyr::select(GEOGRAPHY_CODE, services.share)
```

## Unemployment 2011

```{r}
unemployment <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_1046_1.data.csv?date=latest&geography=1119879169...1119882228,1119882230...1119885230,1119885232...1119885236,1119885238...1119885256,1119885263...1119885265,1119885267,1119885257...1119885262,1119885266,1119885268...1119885792&c_sex=0&c_age=0&c_ecopuk11=0,1,9&measures=20100&signature=NPK-8f824e8ae641543e1e4ae2:0x9e220902be948949eb24ef622423687fba63356a")

unemployment <- unemployment %>%
  dplyr::select(GEOGRAPHY_CODE, GEOGRAPHY_NAME, C_ECOPUK11_NAME, OBS_VALUE) %>% 
  spread(C_ECOPUK11_NAME, OBS_VALUE) %>% 
  mutate(unemployment.share = (`Economically active: Unemployed (including full-time students)` / 
                                 `Economically active: Total`)) %>%
  dplyr::select(GEOGRAPHY_CODE, unemployment.share)
```

## Universities

```{r}
# source: https://www.hesa.ac.uk/support/providers, keep HEI

url <- "https://www.hesa.ac.uk/support/providers"
webpage <- xml2::read_html(url)

uni.table <- rvest::html_table(webpage)[[1]] %>% 
  tibble::as_tibble(.name_repair = "unique") %>%               # repair the repeated columns
  filter(Region!= "Scotland" & Region!="Northern Ireland") %>% # drop Scotland and N. Ireland
  filter(`Provider type` == "HEI")

latlong <-  geocode_OSM(uni.table$Name) # 131 results

# wrong results based on the leaflet map below
# these universities are added in the uni.table.nomatch
wrong.location <- c("University of Plymouth", 
                    "The University of Southampton",
                    "The Open University",
                    "The University of Portsmouth",
                    "Goldsmiths College")

`%nin%` = Negate(`%in%`)

latlong <- latlong %>%
  filter(query %nin% wrong.location)

uni.table.nomatch <- tibble(
  Name = c("Conservatoire for Dance and Drama",
           "Hartpury University",
           "Imperial College of Science, Technology and Medicine",
           "Rose Bruford College of Theatre and Performance",
           "Royal Holloway and Bedford New College",
           "St George's, University of London",
           "The University of Huddersfield",
           "The University of Wales (central functions)",
           "University of London (Institutes and activities)",
           "University of Northumbria at Newcastle",
           "University of St Mark and St John",
           "University of the Arts, London",
           "University of Plymouth", 
           "The University of Southampton",
           "The Open University",
           "The University of Portsmouth",
           "Goldsmiths College"),
  postcode = c("WC1H 9JJ",
               "GL19 3BE",
               "SW7 2BU",
               "DA15 9DF",
               "TW20 0EX",
               "SW17 0RE",
               "HD1 3DH",
               "CF10 3NS",
               "WC1E 7HU",
               "NE1 8QH",
               "PL6 8BH",
               "WC1V 7EY",
               "PL4 8AA",
               "SO17 1BJ",
               "MK7 6BJ",
               "PO1 2UP",
               "SE14 6NW"))

latlong.nomatch <-  geocode_OSM(uni.table.nomatch$postcode) # 16 results
latlong.nomatch$query <- uni.table.nomatch$Name

uni.table <- rbind(latlong, latlong.nomatch)
uni.table <- uni.table %>%
  dplyr::select(query, lat, lon) %>%
  rename(university = query)

# convert uni points to spatial object
coords_uni <- cbind(uni.table$lon, uni.table$lat)
uni.bua <- SpatialPointsDataFrame(coords_uni, data = data.frame(uni.table))
proj4string(uni.bua) <- CRS("+init=epsg:4326") #define projection

# spatial join to bua_buasd
uni.sp <- over(uni.bua, bua_buasd[, "bua11cd"]) # not a spatial object
uni.table$bua11cd <- uni.sp$bua11cd

# Uni map. NOT TO RUN
# leaflet(bua_buasd) %>% 
#   addPolygons() %>% 
#   addCircles(lng=uni.table$lon, lat=uni.table$lat, popup=uni.table$university) %>%
#   addMarkers(lng=uni.table$lon, lat=uni.table$lat, popup=uni.table$university)

uni.freq <- uni.table %>%
  group_by(bua11cd) %>%
  summarise(uni.freq = n())

uni.freq <- merge(bua_buasd@data, uni.freq, by = "bua11cd", all.x = T)
uni.freq$uni.freq <- ifelse(is.na(uni.freq$uni.freq), 0, uni.freq$uni.freq)

uni.freq <- uni.freq %>%
  dplyr::select(bua11cd, uni.freq) %>%
  rename(GEOGRAPHY_CODE = bua11cd)

path.uni <- paste(path, "/data/uk/raw/uni_freq.csv", sep = "")
write_csv(uni.freq, path.uni)
```

## Merge objects

```{r}
data <- list(bb2017, bb2011, pop, pop.dens, ur.ew, british.pop, work.home,
services, unemployment, uni.freq)
sapply(data, function(x) dim(x))
sapply(data, function(x) names(x))

# rename bua11cd to GEOGRAPHY_CODE
names(data[[1]])[1] <- "GEOGRAPHY_CODE"
names(data[[1]])[2] <- "download2017"
names(data[[1]])[3] <- "upload2017"
names(data[[1]])[4] <- "n.tests2017"

names(data[[2]])[1] <- "GEOGRAPHY_CODE"
names(data[[2]])[2] <- "download2011"
names(data[[2]])[3] <- "upload2011"
names(data[[2]])[4] <- "n.tests2011"

names(data[[5]])[1] <- "GEOGRAPHY_CODE"

# merge with reduce
data <- data %>% 
  reduce(inner_join, by = "GEOGRAPHY_CODE")
sapply(data, function(x) sum(is.na(x)))

# NOT RUN
# export file for future reference / backup
# data.out.path <- paste0(path, "/data/data_inter/data_for_uk_regressions.csv")
# write.csv(data, data.out.path)
```

## descriptives

```{r}
ggplot(data = data, aes(x=download2011, y=r_diff_tr)) + geom_point() + geom_smooth(method = "lm")
ggplot(data = data, aes(x=download2017, y=r_diff_tr)) + geom_point() + geom_smooth(method = "lm")

# ggplot(data[(data$n.tests2017>50 & data$download2017<60000),], aes(x=download2017, y=r_diff_tr)) + geom_point() + geom_smooth(method = "lm")
# ggplot(data[(data$n.tests2011>50),], aes(x=download2011, y=r_diff_tr)) + geom_point() + geom_smooth(method = "lm")

f.n.tests2011 <- data %>%
  dplyr::select(n.tests2011) %>%
  group_by(ints = cut_width(n.tests2011, width = 20, boundary = 0)) %>%
  summarise(n = n())
head(f.n.tests2011)
```

## Regressions

F/or all the regressions, we include the aerial units with more than 20 tests in 2011.
This leads to a decrease in the number of observations include in the regressions from 6096 to 6096 - 2561 = 3535

```{r}

base <- lm(r_diff_tr ~ log(download2011) + log(pop2011) + I(n.tests2011/pop2011) + unemployment.share + 
                     british.pop.share +  pop.dens + work.home.share + services.share, data = data, subset = n.tests2011>20)
#nobs(base)

# for robust with library(jtools)
# summ(base, robust = "HC1")

# upload is not a significant predictor

base_glm <- glm(r_diff_tr ~ log(download2011) + log(pop2011) + I(n.tests2011/pop2011) + unemployment.share +
                          british.pop.share +  pop.dens + work.home.share + services.share, data = data, subset = n.tests2011>20)
#nobs(base_glm)

#base_betareg <- betareg(r_diff_tr ~ log(download2011) + log(pop2011) + log(pop2018) + I(n.tests2011/pop2011) + unemployment.share +
#                          british.pop.share +  pop.dens + work.home.share + services.share, data = data, 
#                          link = "logit", subset = n.tests2011>20)
#nobs(base_betareg)

# not converging, try quasipoisson
#base_poisson <- glm(r_diff_tr ~ log(download2011) + log(pop2011) + I(n.tests2011/pop2011) + unemployment.share +
#                          british.pop.share +  pop.dens + work.home.share + services.share, data = data, subset = n.tests2011>20, 
#                          family = "poisson")
#nobs(base_poisson)

# interactions
int_density <- lm(r_diff_tr ~ log(download2011)*pop.dens + log(pop2011) + I(n.tests2011/pop2011) + unemployment.share + 
                     british.pop.share +  pop.dens + work.home.share + services.share, data = data, subset = n.tests2011>20)
#nobs(int_density)

int_pop <- lm(r_diff_tr ~ log(download2011)*log(pop2011) + log(pop2011) + I(n.tests2011/pop2011) + unemployment.share + 
                     british.pop.share +  pop.dens + work.home.share + services.share, data = data, subset = n.tests2011>20)
#nobs(int_pop)

# 2 SLS with IV
model_iv1 <- ivreg(r_diff_tr ~ log(download2011) + log(pop2011) + I(n.tests2011/pop2011) + unemployment.share + 
                     british.pop.share +  pop.dens + work.home.share + services.share 
                   | log(pop2011) + I(n.tests2011/pop2011) + unemployment.share + 
                     british.pop.share +  pop.dens + work.home.share + services.share + uni.freq
                     , data = data, subset = n.tests2011>30)
#summary(model_iv1, vcov = sandwich, diagnostics = TRUE)
#nobs(base)

model_iv2 <- ivreg(r_diff_tr ~ log(download2011) + log(pop2011) + I(n.tests2011/pop2011) + unemployment.share + 
                     british.pop.share +  pop.dens + work.home.share + services.share 
                   | log(pop2011) + I(n.tests2011/pop2011) + unemployment.share + 
                     british.pop.share +  pop.dens + work.home.share + services.share + uni.freq + n.tests2011
                     , data = data, subset = n.tests2011>30)
#summary(model_iv2, vcov = sandwich, diagnostics = TRUE) #yes
```

## Table of Descriptives and Regression tables

```{r}
# Table 9
data.for.decr <- data %>%
  dplyr::select(r_diff_tr, download2011, pop2011, n.tests2011,  
         unemployment.share, british.pop.share, pop.dens, 
         work.home.share, services.share, uni.freq) %>%
  mutate(n.tests2011perhab = n.tests2011/pop2011) %>%
  dplyr::select(-n.tests2011) %>%
  dplyr::select(r_diff_tr, download2011, pop2011, n.tests2011perhab,  # re-redring
         unemployment.share, british.pop.share, pop.dens, 
         work.home.share, services.share, uni.freq)

path.out <- paste(path, "/outputs/table9.htm", sep = "")

stargazer(as.data.frame(data.for.decr), out=path.out, type="html", summary = T,
          covariate.labels=c("normalized difference in ranks, 2011-14", "download speed, 2011", "population, 2011",  
                             "broadband tests per capita, 2011", "% of unemployment, 2011", "% of British population, 2011", 
                             "population density, 2011", "% of people working from home, 2011",
                             "employment in service, 2011 (%)", "Number of universities"))

#Table 10
#regressions table robust SE
rob.base        <- coeftest(base, function(x) vcovHC(x, type="HC0"))
rob.int_density <- coeftest(int_density, function(x) vcovHC(x, type="HC0"))
rob.int_pop <- coeftest(int_pop, function(x) vcovHC(x, type="HC0"))
rob.base_glm <- coeftest(base_glm, function(x) vcovHC(x, type="HC0"))

path.out <- paste(path, "/outputs/table10.htm", sep = "")

stargazer(base, int_density, int_pop, base_glm, type="html",
          se = list(rob.base[,"Std. Error"], rob.int_density[,"Std. Error"], rob.int_pop[,"Std. Error"], rob.base_glm[, "Std. Error"]), 
          dep.var.labels=c("Normalized difference in ranks, 2011-18"),
          #dep.var.caption = "Normalized difference in ranks, 2011-18",
          column.labels = c("OLS", "GLM"),
          column.separate = c(3,1),
          covariate.labels=c("download speed, 2011 (log)", "population, 2011 (log)",  
                             "broadband tests per capita, 2011", "% of unemployment, 2011", "% of British population, 2011", 
                             "population density, 2011", "% of people working from home, 2011",
                             "employment in service, 2011 (%)", "download speed, 2011 (log) x pop. density, 2011", 
                             "download speed, 2011 (log) x population, 2011 (log)"), 
          out=path.out, single.row = FALSE, df = FALSE, omit.stat = c("rsq", "f"), 
          notes = "Robust Std. Error parenthesis")

# Table 11
# 2SLS robust SE table
#model_iv1, model_iv2, 
rob.model_iv1 <- coeftest(model_iv1, function(x) vcovHC(x, type="HC0"))
rob.model_iv2 <- coeftest(model_iv2, function(x) vcovHC(x, type="HC0"))
summ.fit1 <- summary(model_iv1, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)
summ.fit2 <- summary(model_iv2, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)

path.out <- paste(path, "/outputs/table11.htm", sep = "")

stargazer(model_iv1, model_iv2, type = "html", 
          se = list(rob.model_iv1[,"Std. Error"], rob.model_iv2[,"Std. Error"]), 
          dep.var.labels=c("Normalized difference in ranks, 2011-18"),
          covariate.labels=c("download speed, 2011 (log)", "population, 2011 (log)",  
                             "broadband tests per capita, 2011", "% of unemployment, 2011", "% of British population, 2011", 
                             "population density, 2011", "% of people working from home, 2011",
                             "employment in service, 2011 (%)"),
          out=path.out, single.row = FALSE, df = FALSE, 
          omit.stat = c("rsq", "f"), notes = c("Robust Std. Error in parenthesis", "IVs: (1) Number of universities",
                                               "IVs: (2) Number of universities, Number of broadband tests, 2011"),
          add.lines = list(c(rownames(summ.fit1$diagnostics)[1],                    # use rownames as line name. in this case Weak insterumetns
                             round(summ.fit1$diagnostics[1, "statistic"], 2), 
                             round(summ.fit2$diagnostics[1, "statistic"], 2)), 
                           c(rownames(summ.fit1$diagnostics)[2], 
                             round(summ.fit1$diagnostics[2, "statistic"], 2), 
                             round(summ.fit2$diagnostics[2, "statistic"], 2)), 
                           c("P-value", 
                             round(summ.fit1$diagnostics[2, "p-value"], 2), 
                             round(summ.fit2$diagnostics[2, "p-value"], 2)), 
                           c(rownames(summ.fit1$diagnostics)[3], 
                             round(summ.fit1$diagnostics[3, "statistic"], 2), 
                             round(summ.fit2$diagnostics[3, "statistic"], 2)),
                           c("P-value", 
                             round(summ.fit1$diagnostics[3, "p-value"], 2), 
                             round(summ.fit2$diagnostics[3, "p-value"], 2))
          ))
```

## Autocorrelation

The below indicates no Spatial Autocorrelation in r_diff_tr and in the residulas of the base regression.

```{r}
# auto correlation

# create a copy of the spatial object
bua_buasd_copy <- bua_buasd

# merge data used in the regressions with the spatial object 
bua_buasd_copy@data <- data.frame(bua_buasd_copy@data, data[match(bua_buasd_copy@data$bua11cd, data$GEOGRAPHY_CODE),])
bua_buasd_copy <- sp.na.omit(bua_buasd_copy)

# Moran's I
nearest_neigh = knearneigh(coordinates(bua_buasd_copy),k=6, RANN=F)
nearest_neigh = knn2nb(nearest_neigh)
spweights = nb2listw(nearest_neigh)

# Moran's I plot
moran.plot(data$r_diff_tr, spweights)

# Moran's I test
moran.test(data$r_diff_tr, spweights)

# subset bua_buasd based on the obs. included in base
inc.obs <- augment(base)[1]
data.inc.obs <-data %>% 
  slice(as.numeric(inc.obs$.rownames))

# merge data used in the base regression with the spatial object 
bua_buasd_copy <- bua_buasd
bua_buasd_copy@data <- data.frame(bua_buasd_copy@data, data[match(bua_buasd_copy@data$bua11cd, data.inc.obs$GEOGRAPHY_CODE),])
bua_buasd_copy <- sp.na.omit(bua_buasd_copy)

nearest_neigh = knearneigh(coordinates(bua_buasd_copy),k=6, RANN=F)
nearest_neigh = knn2nb(nearest_neigh)
spweights = nb2listw(nearest_neigh)

# Autocorrelation in residuals: NO AUTOCORRELATION
lm.morantest(base, spweights)
```

```{r}
path.out <- paste0(path, "/data/data_inter/UK.RData")
save.image(path.out)
```