### 2020-09-02 src/un.Rmd

---
title: "UN"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../outputs")
  })
---

```{r}
library(rmarkdown)
library(data.table)
library(tidyverse)
library(readxl)
library(httr)
library(rprojroot)

library(ggplot2)
#library(geojsonio)
library(rgdal)
library(maptools)
library(sp)
library(sf)
library(rgeos)
#library(classInt)
#library(RColorBrewer)
library(knitr)
#library(viridis)
#library(hrbrthemes)
library(fs)
library(betareg)
library(broom)

library(AER)
library(MASS)
#library(foreign)
library(stargazer)
library(ivpack) 
library(leaflet)
library(rvest)
library(xml2)
library(tmaptools)
library(spdep)
library(spatialreg)

library(acs)
library(tidycensus)
library(rio)
library(wbstats)
library(WDI)
library(kableExtra)

# This is the project path
path <- find_rstudio_root_file()
```

## Zipf regressions

```{r}
# get urban pop data
url <- "https://population.un.org/wup/Download/Files/WUP2018-F22-Cities_Over_300K_Annual.xls"
GET(url, write_disk(tf <- tempfile(fileext = ".xls")))
urban.pop.raw <- read_excel(tf, 1L, range = "A17:CP1877", col_names = T)
# also saved as ./Data/UN/WUP2018-F22-Cities_Over_300K_Annual.xls

# get country codes
url <- "https://population.un.org/wup/Download/Files/WUP2018-F00-Locations.xlsx"
# also saved as ./Data/UN/WUP2018-F22-Cities_Over_300K_Annual.xls

codes <- rio::import(url)  # , startRow = 32)
codes <- codes[27:285,2:5]
codes <- codes %>%
  dplyr::select(-`...3`) %>%
  rename(`Country or area` = `...2`,
         `Country Code` = `United Nations`,
         ISO = `...5`) %>%
  remove_rownames() %>%
  dplyr::select(-`Country or area`) # no need of this at the end

# names
names <- urban.pop.raw %>% 
  dplyr::select(1:5 | starts_with("2")) %>%
  dplyr::select(-(25:41)) %>%
  rename_at(vars(-c(Index, 
                    `Country Code`, 
                    `Country or area`, 
                    `City Code`, 
                    `Urban Agglomeration`)),
            function(x) paste0("pop", x))

# # rankings OLD
# ranks <- urban.pop.raw %>% 
#   dplyr::select(`Country Code` | starts_with("2")) %>%
#   group_by(`Country Code`) %>%
#   summarise(across(everything(),dense_rank)) %>%
#   dplyr::select(-c(1, 21:37)) %>%
#   rename_at(vars(-`Country Code`),
#             function(x) paste0("rank", x))

dense_rank_d <- function(x){
  dense_rank(-x)
}

# rankings
ranks <- urban.pop.raw %>% 
  dplyr::select(`Country Code`, `City Code`, starts_with("2")) %>%
  group_by(`Country Code`) %>%
  mutate_at(3:38, dense_rank_d) %>%
  #mutate(dense_rank(-c(3:38))) %>%
  dplyr::select(-c(22:38)) %>%
  rename_at(vars(-`Country Code`, -`City Code`),
            function(x) paste0("rank", x))

# merge
#urban.pop <- bind_cols(names, ranks)
urban.pop <- left_join(names, ranks, by = "City Code") %>%
  dplyr::select(-`Country Code.y`) %>%
  rename(`Country Code` = `Country Code.x`)

#urban.pop$`Country Code...25` <- NULL
#urban.pop <- rename(urban.pop, `Country Code` = `Country Code...2`)
urban.pop <- merge(urban.pop, codes, by = "Country Code", all.x = T)
urban.pop <- urban.pop %>%
  relocate(ISO, .before = `City Code`) %>%
  dplyr::select(-Index) %>%
  group_by(ISO) %>%
  add_tally %>%
  rename(n.cities = n) %>%
  relocate(n.cities, .before = pop2000) %>%
  filter(n.cities > 5) # keep the countries with x or more urban agglomeration of 300k or above

# Zipf regressions
zipf <- NULL
years <- 2000:2018
countries <- unique(urban.pop$ISO)
for (c in countries){
  for (i in years){
  model.name <- paste(c, i, sep = ".")
  formula <- paste0("log(rank", i, " - 0.5) ~ log(pop", i, ")") # log transformation and Gabaix and Ibragimov (2011) correction
  model <- lm(formula, data = urban.pop[urban.pop$ISO==c,])
  b <- coef(summary(model))[2, 1]
  se <- coef(summary(model))[2, 2]
  df <- cbind(b = b, se = se, country = c, year = i)
  zipf <- rbind(zipf, df)
  # assign(model.name, model) # assign object (x) to a name (model.name)
}
}

# model$coefficients[2]
# se <- sqrt(diag(vcov(model)))
# coef(summary(model))[2, 2]

zipf <- as_tibble(zipf) %>% 
  mutate(b = as.numeric(b),
         se = as.numeric(se),
         year = as.numeric(year)) %>%
  dplyr::select(country, year, b, se)
```

## Data for second stage regressions - search World Bank API

```{r}

# This are just asearches in the world bank database

wb_search(pattern = "broadband")
wb_search(pattern = "internet")
wb_search(pattern = "mobile")
wb_search(pattern = "fixed")
wb_search(pattern = "total population")
wb_search(pattern = "population density")
wb_search(pattern = "GDP per capita")
wb_search(pattern = "GDP growth")
wb_search(pattern = "trade")
wb_search(pattern = "government final consumption expenditure")
wb_search(pattern = "value added")
wb_search(pattern = "Mortality rate")
wb_search(pattern = "Female labor force")
wb_search(pattern = "Secure internet servers")
wb_search(pattern = "households with television")
wb_search(pattern = "telecommunication sector")

# IT.NET.BBND.P3	    Broadband subscribers (per 1,000 people)
# IT.NET.USER.P3	    Internet users (per 1,000 people)	
# IT.CEL.SETS.P3      Mobile phone subscribers (per 1,000 people)	
# IT.MLT.MAIN.P3	    Telephone mainlines (per 1,000 people)	
# SP.POP.TOTL         Population, total	
# EN.POP.DNST	        Population density (people per sq. km of land area)
# NY.GDP.PCAP.PP.KD	  GDP per capita, PPP (constant 2017 international $)
# NY.GDP.MKTP.KD.ZG	  GDP growth (annual %)
# NE.TRD.GNFS.ZS	    Trade (% of GDP)
# NE.CON.GOVT.ZS	    General government final consumption expenditure (% of GDP)
# NV.AGR.TOTL.ZS	    Agriculture, forestry, and fishing, value added (% of GDP)
# HF.DYN.MORT	        Mortality rate, under-5 (per 1,000)	
# SL.TLF.TOTL.FE.ZS	  Labor force, female (% of total labor force)	
# IT.NET.SECR.P6	    Secure Internet servers (per 1 million people)
# IT.TVS.HOUS.ZS	    Households with television (%) **20 YEAR LAG IS NEEDED**
# SL.EMP.TRAN	        Number of people employed in transportation and telecommunication sector **20 YEAR LAG IS NEEDED**

# test <- wb_data(indicator = "IT.NET.BBND.P3", start_date = 2000, end_date = 2018)

wb_data("IT.NET.BBND.P3") # no, Broadband subscribers (per 1,000 people)
wb_data("IT.NET.BBND")    # Broadband subscribers
WDI("IT.NET.BBND.P3", country = "all") # until 2012
# localy saved in ./Data/UN/API_IT.NET.BBND.P2_DS2_en_csv_v2_1221909.zip

wb_data("IT.NET.USER.P3") # no, Internet users (per 1,000 people)	
wb_data("IT.NET.USER")    # no, Internet users (per 1,000 people)	
# http://api.worldbank.org/v2/country/all/indicator/IT.NET.USER.P3?per_page=4000
WDI("IT.NET.USER.P3", country = "all") # until 2012
# localy saved in ./Data/UN/API_IT.NET.USER.ZS_DS2_en_csv_v2_1307380.zip

wb_data("IT.CEL.SETS.P3") # no, Mobile phone subscribers (per 1,000 people)	
wb_data("IT.CEL.SETS")    # Mobile phone subscribers
WDI("IT.CEL.SETS.P3", country = "all") # until 2012
# localy saved in ./Data/UN/API_IT.CEL.SETS.P2_DS2_en_csv_v2_1217961.zip

wb_data("IT.MLT.MAIN.P3") # no, Telephone mainlines (per 1,000 people)	
wb_data("IT.MLT.MAIN")    # Telephone mainlines 
# localy saved in ./Data/UN/API_IT.MLT.MAIN.P2_DS2_en_csv_v2_1221907.zip

wb_data("SP.POP.TOTL")
wb_data("EN.POP.DNST")
wb_data("NY.GDP.PCAP.PP.KD")
wb_data("NY.GDP.MKTP.KD.ZG")
wb_data("NE.TRD.GNFS.ZS")
wb_data("NE.CON.GOVT.ZS")
wb_data("NV.AGR.TOTL.ZS")
wb_data("HF.DYN.MORT")
wb_data("SL.TLF.TOTL.FE.ZS")
wb_data("IT.NET.SECR.P6") # no, too many missing
wb_data("IT.TVS.HOUS.ZS") # no, Households with television (%) **20 YEAR LAG IS NEEDED**
WDI("IT.TVS.HOUS.ZS", country = "all") #no 
# localy saved in ./Data/UN/
wb_data("SL.EMP.TRAN")    #	no, Number of people employed in transportation and telecommunication sector **20 YEAR LAG IS NEEDED**
# localy saved in ./Data/UN/
WDI("SL.EMP.TRAN", country = "all") #no
```

## Data for second stage regressions

```{r}
# Getting the data for the second stage (after Zipf) regressions.
# This is where the data build starts.
# The below codes were obtained after the above searches.

rhs1 <- wb_data(c("SP.POP.TOTL",
                  "EN.POP.DNST",
                  "NY.GDP.PCAP.PP.KD",
                  "NY.GDP.MKTP.KD.ZG",
                  "NE.TRD.GNFS.ZS",
                  "NE.CON.GOVT.ZS",
                  "NV.AGR.TOTL.ZS",
                  #"HF.DYN.MORT", # too many missing data points
                  #"IT.NET.SECR.P6", # too many missing data points
                  "SL.TLF.TOTL.FE.ZS"),
                start_date = 2000,
                end_date = 2018)

rhs1 <- rhs1 %>% 
  rename(country_name = country,
         country = iso3c,
         year = date,
         pop.dens = EN.POP.DNST,
         # mort5 = HF.DYN.MORT,
         # sec.serv = IT.NET.SECR.P6,  
         gov.exp = NE.CON.GOVT.ZS,
         trade = NE.TRD.GNFS.ZS,
         agri = NV.AGR.TOTL.ZS,
         gdp.growth = NY.GDP.MKTP.KD.ZG, 
         gdppc = NY.GDP.PCAP.PP.KD,
         fem.lab = SL.TLF.TOTL.FE.ZS,
         pop = SP.POP.TOTL) %>%
  mutate(non.agri = 100-agri) %>%
  dplyr::select(-agri)

# internet users
net.path <- paste0(path, "/Data/UN/API_IT.NET.USER.ZS_DS2_en_csv_v2_1307380.zip")
con <- unz(net.path, "API_IT.NET.USER.ZS_DS2_en_csv_v2_1307380.csv")
net <- read_csv(con, skip = 4) %>%
  dplyr::select(-(5:44), -(64:65)) %>%
  rename(country = `Country Code`) %>% 
  dplyr::select(-c(1,3:4)) %>%
  gather("year", "net100hab", -c(country)) %>%
  mutate(year = as.numeric(year))

# broadband users
bb.path <- paste0(path, "/Data/UN/API_IT.NET.BBND.P2_DS2_en_csv_v2_1221909.zip")
con <- unz(bb.path, "API_IT.NET.BBND.P2_DS2_en_csv_v2_1221909.csv")
bb <- read_csv(con, skip = 4) %>%
  dplyr::select(-(5:44), -(64:65)) %>%
  rename(country = `Country Code`) %>% 
  dplyr::select(-c(1,3:4)) %>%
  gather("year", "bb100hab", -c(country)) %>%
  mutate(year = as.numeric(year))

# mobile phone users
mobile.path <- paste0(path, "/Data/UN/API_IT.CEL.SETS.P2_DS2_en_csv_v2_1217961.zip")
con <- unz(mobile.path, "API_IT.CEL.SETS.P2_DS2_en_csv_v2_1217961.csv")
mobile <- read_csv(con, skip = 4) %>%
  dplyr::select(-(5:44), -(64:65)) %>%
  rename(country = `Country Code`) %>% 
  dplyr::select(-c(1,3:4)) %>%
  gather("year", "mobile100hab", -c(country)) %>%
  mutate(year = as.numeric(year))

# fixed telephony users
fixed.path <- paste0(path, "/Data/UN/API_IT.MLT.MAIN.P2_DS2_en_csv_v2_1221907.zip")
con <- unz(fixed.path, "API_IT.MLT.MAIN.P2_DS2_en_csv_v2_1221907.csv")
fixed <- read_csv(con, skip = 4) %>%
  dplyr::select(-(5:44), -(64:65)) %>%
  rename(country = `Country Code`) %>% 
  dplyr::select(-c(1,3:4)) %>%
  gather("year", "fixed100hab", -c(country)) %>%
  mutate(year = as.numeric(year))

# tv 
# tv <- WDI("IT.TVS.HOUS.ZS", country = "all") %>%
#   rename(country = iso2c, country_name = country, tv.hous100 = IT.TVS.HOUS.ZS) %>% 
#   mutate(year = year + 10) # this is for the 10 year lag IV

tv.path <- paste0(path, "/Data/UN/itu_database2.xls")
tv <- read_excel(tv.path) %>%
  filter(`Series Name` == "Percentage of households with TV") %>%
  dplyr::select(-(5:35)) %>% #, -(64:65)
  rename(country = `Country Code`) %>% 
  dplyr::select(-c(1,2,4)) %>%
  gather("year", "tv100hh", -c(country)) %>%
  mutate(tv100hh = as.numeric(tv100hh),
         year = as.numeric(year) + 10) %>% # this is for the 10 year lag IV 
  rename(tv100hh10L = tv100hh)

# employment in telecommunications
tele.emp.path <- paste0(path, "/Data/UN/tele_employment.csv")
tele.emp <- read_csv(tele.emp.path) %>%
  dplyr::select(-(5:34)) %>% #, -(64:65)
  rename(country = `Country Code`) %>% 
  dplyr::select(-c(1,2,4)) %>%
  gather("year", "tele.emp", -c(country)) %>%
  mutate(tele.emp = as.numeric(tele.emp),
         year = as.numeric(year) + 10) %>% # this is for the 10 year lag IV 
  rename(tele.emp10L = tele.emp)

# TODO: GAPS
```

## Merge objects

```{r}
data <- list(zipf, rhs1, net, bb, mobile, fixed)# tele.emp and tv are not used in the regressions
sapply(data, function(x) dim(x))
sapply(data, function(x) names(x))
sapply(data, function(x) str(x))

# merge with reduce
data <- data %>% 
  reduce(inner_join, by = c("country", "year")) %>%
  relocate(iso2c, .after = country) %>%
  relocate(country_name, .after = country)
sapply(data, function(x) sum(is.na(x)))

# NOT RUN
# export file for future reference / backup
# data.out.path <- paste0(path, "/data/data_inter/data_for_un_regressions.csv")
# write.csv(data, data.out.path)
```

## descriptives

```{r}
ggplot(data = data, aes(x=bb100hab, y=b)) + geom_point() + geom_smooth(method = "lm")
```

## regressions

```{r}

# These are just tests

library(plm)
data.plm <- pdata.frame(data, index=c("country","year"), drop.index=TRUE, row.names=TRUE)

test.lm <- plm(b ~ pop.dens  + gov.exp + trade + non.agri + gdp.growth + gdppc + fem.lab + pop + net100hab + as.factor(year), 
               data = data, model = "within")
summary(test.lm)

test.2w <- plm(b ~ fixed100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop), 
               data = data, model = "within", effect = "twoways", weights = 1/(se^2))
summary(test.2w)

test.w.ols <- lm(b ~ net100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop)
                  + as.factor(country) + as.factor(year), 
               data = data, weights = 1/(se^2))
summary(test.w.ols)

test.2w.2sls <- ivreg(b ~ fixed100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country) |
                      log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country)  +
                      fem.lab, 
               data = data, weights = 1/(se^2))
summary(test.2w.2sls, vcov = sandwich, diagnostics = TRUE)

test.2w.2sls <- plm(b ~ fixed100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country) |
                      log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country)  +
                      fem.lab, 
               data = data, weights = 1/(se^2), model = "fd", effect = "twoways")
summary(test.2w.2sls, vcov = sandwich, diagnostics = TRUE)

# + bb100hab + mobile100hab + fixed100hab 
# + mort5 + fem.lab +

```

## Table 4

```{r}
net.model <- lm(b ~ net100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop)
                  + as.factor(country) + as.factor(year), 
               data = data, weights = 1/(se^2))

bb.model <- lm(b ~ bb100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop)
                  + as.factor(country) + as.factor(year), 
               data = data, weights = 1/(se^2))

fixed.model <- lm(b ~ fixed100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop)
                  + as.factor(country) + as.factor(year), 
               data = data, weights = 1/(se^2))

mobile.model <- lm(b ~ mobile100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop)
                  + as.factor(country) + as.factor(year), 
               data = data, weights = 1/(se^2))

# Table 4
net.rse <- coeftest(net.model, function(x) vcovHC(x, type="HC0"))
bb.rse <- coeftest(bb.model, function(x) vcovHC(x, type="HC0"))
fixed.rse <- coeftest(fixed.model, function(x) vcovHC(x, type="HC0"))
mobile.rse <- coeftest(mobile.model, function(x) vcovHC(x, type="HC0"))

path.out <- paste(path, "/outputs/table4.htm", sep = "")

stargazer(net.model, bb.model, fixed.model, mobile.model, type="html",
          se = list(net.rse[,"Std. Error"], bb.rse[,"Std. Error"], fixed.rse[,"Std. Error"], mobile.rse[, "Std. Error"]), 
          dep.var.labels=c("Zipf Coefficient 2000-18"),
          # covariate.labels=c("households w. internet, 2013 (%)", 
          #                    "population, 2013",  
          #                    "% of unemployment, 2013", 
          #                    "% of white population, 2013", 
          #                    "income, 2013 (log)",
          #                    "population density, 2013", 
          #                    "employment in service, 2011 (%)", 
          #                    "commute in minutes, 2013", 
          #                    "households w. internet, 2013 (%) x pop. density, 2013", 
          #                    "households w. internet, 2013 (%) x population, 2013"), 
          out=path.out, single.row = FALSE, df = FALSE, 
          omit = c("country", "year"),
          add.lines = list(c("Country fixed effects", "Yes", "Yes", "Yes", "Yes"),
                           c("Yearly fixed effects", "Yes", "Yes", "Yes", "Yes")),
          omit.stat = c("rsq", "f"), notes = "Robust Std. Error in parenthesis, corrected Zipf coefficient following Gabaix and Ibragimov (2011)")

```

## Table 5

```{r}
net.2sls <- ivreg(b ~ net100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country) |
                      log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country)  +
                      fem.lab, 
               data = data, weights = 1/(se^2))
summary(net.2sls, vcov = sandwich, diagnostics = TRUE)

# net.2sls.2 <- ivreg(b ~ net100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
#                       + as.factor(year) + as.factor(country) |
#                       log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
#                       + as.factor(year) + as.factor(country)  +
#                       fem.lab + tv100hh10L, # 
#                data = data, weights = 1/(se^2))
# summary(net.2sls.2, vcov = sandwich, diagnostics = TRUE)

bb.2sls <- ivreg(b ~ bb100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country) |
                      log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country)  +
                      fem.lab, 
               data = data, weights = 1/(se^2))
summary(bb.2sls, vcov = sandwich, diagnostics = TRUE)

fixed.2sls <- ivreg(b ~ fixed100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country) |
                      log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country)  +
                      fem.lab, 
               data = data, weights = 1/(se^2))
summary(fixed.2sls, vcov = sandwich, diagnostics = TRUE)

mobile.2sls <- ivreg(b ~ mobile100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country) |
                      log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country)  +
                      fem.lab, 
               data = data, weights = 1/(se^2))
summary(mobile.2sls, vcov = sandwich, diagnostics = TRUE)

net.2sls.rse <- coeftest(net.2sls, function(x) vcovHC(x, type="HC0"))
summ.net.2sls <- summary(net.2sls, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)

bb.2sls.rse <- coeftest(bb.2sls, function(x) vcovHC(x, type="HC0"))
summ.bb.2sls <- summary(bb.2sls, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)

fixed.2sls.rse <- coeftest(fixed.2sls, function(x) vcovHC(x, type="HC0"))
summ.fixed.2sls <- summary(fixed.2sls, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)

mobile.2sls.rse <- coeftest(mobile.2sls, function(x) vcovHC(x, type="HC0"))
summ.mobile.2sls <- summary(mobile.2sls, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)

path.out <- paste(path, "/outputs/table5.htm", sep = "")

stargazer(net.2sls, bb.2sls, fixed.2sls, mobile.2sls, type = "html", 
          se = list(net.2sls.rse[,"Std. Error"], bb.2sls.rse[,"Std. Error"],
                    fixed.2sls.rse[,"Std. Error"], mobile.2sls.rse[,"Std. Error"]), 
          dep.var.labels=c("Normalized difference in ranks, 2013-18"),
          # covariate.labels=c("households w. internet, 2013 (%)", 
          #                    "population, 2013",  
          #                    "% of unemployment, 2013", 
          #                    "% of white population, 2013", 
          #                    "income, 2013 (log)",
          #                    "population density, 2013", 
          #                    "employment in service, 2011 (%)", 
          #                    "commute in minutes, 2013"), 
          out=path.out, single.row = FALSE, df = FALSE,
          omit = c("country", "year"),
          omit.stat = c("rsq", "f"), notes = c("Robust Std. Error in parenthesis", 
                                               "IV: Female paticipation in labour force"),
          add.lines = list(c(rownames(summ.net.2sls$diagnostics)[1],                    # use rownames as line name. in this case Weak insterumetns
                             round(summ.net.2sls$diagnostics[1, "statistic"], 2), 
                             round(summ.bb.2sls$diagnostics[1, "statistic"], 2),
                             round(summ.fixed.2sls$diagnostics[1, "statistic"], 2),
                             round(summ.mobile.2sls$diagnostics[1, "statistic"], 2)), 
                           c(rownames(summ.net.2sls$diagnostics)[2], 
                             round(summ.net.2sls$diagnostics[2, "statistic"], 2), 
                             round(summ.bb.2sls$diagnostics[2, "statistic"], 2),
                             round(summ.fixed.2sls$diagnostics[2, "statistic"], 2),
                             round(summ.mobile.2sls$diagnostics[2, "statistic"], 2)), 
                           c("P-value", 
                             round(summ.net.2sls$diagnostics[2, "p-value"], 2), 
                             round(summ.bb.2sls$diagnostics[2, "p-value"], 2),
                             round(summ.fixed.2sls$diagnostics[2, "p-value"], 2),
                             round(summ.mobile.2sls$diagnostics[2, "p-value"], 2)), 
                           c(rownames(summ.net.2sls$diagnostics)[3], 
                             round(summ.net.2sls$diagnostics[3, "statistic"], 2), 
                             round(summ.bb.2sls$diagnostics[3, "statistic"], 2),
                             round(summ.fixed.2sls$diagnostics[3, "statistic"], 2),
                             round(summ.mobile.2sls$diagnostics[3, "statistic"], 2)),
                           c("P-value", 
                             round(summ.net.2sls$diagnostics[3, "p-value"], 2), 
                             round(summ.bb.2sls$diagnostics[3, "p-value"], 2),
                             round(summ.fixed.2sls$diagnostics[3, "p-value"], 2),
                             round(summ.mobile.2sls$diagnostics[3, "p-value"], 2)),
                           c("Country fixed effects", "Yes", "Yes", "Yes", "Yes"),
                           c("Yearly fixed effects", "Yes", "Yes", "Yes", "Yes")
          ))

```

## Table 1

```{r}
zipf.table <- net.model$model %>%
  dplyr::select(`as.factor(country)`, `as.factor(year)`, b) %>%
  spread(`as.factor(year)`, b) %>%
  rename(`Zipf coef.` = `as.factor(country)`)

zipf.table <- zipf.table %>% kable(format = "html", digits = 3)

path.out <- paste(path, "/outputs/table1.htm", sep = "")
save_kable(zipf.table, path.out)
```

```{r}
path.out <- paste0(path, "/data/data_inter/UN.RData")
save.image(path.out)
```