### 2020-11-18 src/s1.Rmd
S1

---
title: "S1 Appendix"
author: ""
output: pdf_document
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../outputs")
  })
---

```{r, echo=FALSE, warning=FALSE, results='hide', message=FALSE}

# This .Rmd create teh S1 Appendix for the estimation of eq. 3
# using alternative specifications: Gini, Herfindahl-Hirschman
# and CV. 
# The output S1.pdf is located in /outputs

library(rmarkdown)
library(data.table)
library(tidyverse)
library(readxl)
library(httr)
library(rprojroot)

library(ggplot2)
#library(geojsonio)
library(rgdal)
library(maptools)
library(sp)
library(sf)
library(rgeos)
#library(classInt)
#library(RColorBrewer)
library(knitr)
#library(viridis)
#library(hrbrthemes)
library(fs)
library(betareg)
library(broom)

library(AER)
library(MASS)
#library(foreign)
library(stargazer)
library(ivpack) 
library(leaflet)
library(rvest)
library(xml2)
library(tmaptools)
library(spdep)
library(spatialreg)

library(acs)
library(tidycensus)
library(rio)
library(wbstats)
library(WDI)
library(kableExtra)

library(REAT)

# This is the project path
# path <- find_rstudio_root_file()

load("C:/Users/nw19521/OneDrive - University of Bristol/projects/Ioannides/ict.un.us.uk/data/data_inter/UN.RData")
```

```{r, echo=FALSE, warning=FALSE, results='hide', message=FALSE}

# Gini

net.gini.2sls <- ivreg(gini ~ net100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country) |
                      log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country)  +
                      fem.lab, 
               data = data, weights = 1/(se^2))
summary(net.gini.2sls, vcov = sandwich, diagnostics = TRUE)

bb.gini.2sls <- ivreg(gini ~ bb100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country) |
                      log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country)  +
                      fem.lab, 
               data = data, weights = 1/(se^2))
summary(bb.gini.2sls, vcov = sandwich, diagnostics = TRUE)

fixed.gini.2sls <- ivreg(gini ~ fixed100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country) |
                      log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country)  +
                      fem.lab, 
               data = data, weights = 1/(se^2))
summary(fixed.gini.2sls, vcov = sandwich, diagnostics = TRUE)

mobile.gini.2sls <- ivreg(gini ~ mobile100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country) |
                      log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country)  +
                      fem.lab, 
               data = data, weights = 1/(se^2))
summary(mobile.gini.2sls, vcov = sandwich, diagnostics = TRUE)

net.gini.2sls.rse <- coeftest(net.gini.2sls, function(x) vcovHC(x, type="HC0"))
summ.net.gini.2sls <- summary(net.gini.2sls, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)

bb.gini.2sls.rse <- coeftest(bb.gini.2sls, function(x) vcovHC(x, type="HC0"))
summ.bb.gini.2sls <- summary(bb.gini.2sls, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)

fixed.gini.2sls.rse <- coeftest(fixed.gini.2sls, function(x) vcovHC(x, type="HC0"))
summ.fixed.gini.2sls <- summary(fixed.gini.2sls, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)

mobile.gini.2sls.rse <- coeftest(mobile.gini.2sls, function(x) vcovHC(x, type="HC0"))
summ.mobile.gini.2sls <- summary(mobile.gini.2sls, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)
```

```{r, echo=FALSE, warning=FALSE, results='asis'}
stargazer(net.gini.2sls, bb.gini.2sls, fixed.gini.2sls, mobile.gini.2sls, type = "latex",
          se = list(net.gini.2sls.rse[,"Std. Error"], bb.gini.2sls.rse[,"Std. Error"],
                    fixed.gini.2sls.rse[,"Std. Error"], mobile.gini.2sls.rse[,"Std. Error"]),
          header = FALSE,
          no.space=TRUE,
          title = "2SLS estimation of equation (3), alt. specifications (i)\\label{2sls.gini.global}",
          font.size = "small",
          column.sep.width = "1pt",
          dep.var.labels=c("Gini coefficient 2000-18"),
          column.separate = c(1,1,1,1),
          covariate.labels=c("Internet users per 100 hab. (log)",
                             "Broadband users per 100 hab. (log)",
                             "Mobile phone users per 100 hab. (log)",
                             "Fixed phone users per 100 hab. (log)",
                             "Population density (log)",
                             "Government expenditure (\\% GDP)",
                             "Trade (\\% of GDP)",
                             "Non agriculture value added (\\% GDP)",
                             "GDP growth",
                             "GDP per capita (log)",
                             "Population (log)"),
          single.row = FALSE, df = FALSE,
          omit = c("country", "year"),
          omit.stat = c("rsq", "f"),
          add.lines = list(c(rownames(summ.net.gini.2sls$diagnostics)[1], # use rownames as line name. in this case Weak insterumetns
                             round(summ.net.gini.2sls$diagnostics[1, "statistic"], 2),
                             round(summ.bb.gini.2sls$diagnostics[1, "statistic"], 2),
                             round(summ.fixed.gini.2sls$diagnostics[1, "statistic"], 2),
                             round(summ.mobile.gini.2sls$diagnostics[1, "statistic"], 2)),
                           c(rownames(summ.net.gini.2sls$diagnostics)[2],
                             round(summ.net.gini.2sls$diagnostics[2, "statistic"], 2),
                             round(summ.bb.gini.2sls$diagnostics[2, "statistic"], 2),
                             round(summ.fixed.gini.2sls$diagnostics[2, "statistic"], 2),
                             round(summ.mobile.gini.2sls$diagnostics[2, "statistic"], 2)),
                           c("P-value",
                             round(summ.net.gini.2sls$diagnostics[2, "p-value"], 2),
                             round(summ.bb.gini.2sls$diagnostics[2, "p-value"], 2),
                             round(summ.fixed.gini.2sls$diagnostics[2, "p-value"], 2),
                             round(summ.mobile.gini.2sls$diagnostics[2, "p-value"], 2)),
                           # c(rownames(summ.net.2sls$diagnostics)[3],
                           #   round(summ.net.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.bb.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.fixed.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.mobile.2sls$diagnostics[3, "statistic"], 2)),
                           # c("P-value",
                           #   round(summ.net.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.bb.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.fixed.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.mobile.2sls$diagnostics[3, "p-value"], 2)),
                           c("Country fixed effects", "Yes", "Yes", "Yes", "Yes"),
                           c("Yearly fixed effects", "Yes", "Yes", "Yes", "Yes")),
          notes = c("Robust Std. Errors in parenthesis",
                    "IV: Female participation in labour force"))

```

```{r, echo=FALSE, warning=FALSE, results='hide', message=FALSE}

# herf

net.herf.2sls <- ivreg(herf ~ net100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country) |
                      log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country)  +
                      fem.lab, 
               data = data, weights = 1/(se^2))
summary(net.herf.2sls, vcov = sandwich, diagnostics = TRUE)

bb.herf.2sls <- ivreg(herf ~ bb100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country) |
                      log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country)  +
                      fem.lab, 
               data = data, weights = 1/(se^2))
summary(bb.herf.2sls, vcov = sandwich, diagnostics = TRUE)

fixed.herf.2sls <- ivreg(herf ~ fixed100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country) |
                      log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country)  +
                      fem.lab, 
               data = data, weights = 1/(se^2))
summary(fixed.herf.2sls, vcov = sandwich, diagnostics = TRUE)

mobile.herf.2sls <- ivreg(herf ~ mobile100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country) |
                      log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country)  +
                      fem.lab, 
               data = data, weights = 1/(se^2))
summary(mobile.herf.2sls, vcov = sandwich, diagnostics = TRUE)

net.herf.2sls.rse <- coeftest(net.herf.2sls, function(x) vcovHC(x, type="HC0"))
summ.net.herf.2sls <- summary(net.herf.2sls, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)

bb.herf.2sls.rse <- coeftest(bb.herf.2sls, function(x) vcovHC(x, type="HC0"))
summ.bb.herf.2sls <- summary(bb.herf.2sls, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)

fixed.herf.2sls.rse <- coeftest(fixed.herf.2sls, function(x) vcovHC(x, type="HC0"))
summ.fixed.herf.2sls <- summary(fixed.herf.2sls, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)

mobile.herf.2sls.rse <- coeftest(mobile.herf.2sls, function(x) vcovHC(x, type="HC0"))
summ.mobile.herf.2sls <- summary(mobile.herf.2sls, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)
```

```{r, echo=FALSE, warning=FALSE, results='asis'}
stargazer(net.herf.2sls, bb.herf.2sls, fixed.herf.2sls, mobile.herf.2sls, type = "latex",
          se = list(net.herf.2sls.rse[,"Std. Error"], bb.herf.2sls.rse[,"Std. Error"],
                    fixed.herf.2sls.rse[,"Std. Error"], mobile.herf.2sls.rse[,"Std. Error"]),
          header = FALSE,
          no.space=TRUE,
          title = "2SLS estimation of equation (3), alt. specifications (ii)\\label{2sls.herf.global}",
          font.size = "small",
          column.sep.width = "1pt",
          dep.var.labels=c("Herfindahl-Hirschman coefficient 2000-18"),
          column.separate = c(1,1,1,1),
          covariate.labels=c("Internet users per 100 hab. (log)",
                             "Broadband users per 100 hab. (log)",
                             "Mobile phone users per 100 hab. (log)",
                             "Fixed phone users per 100 hab. (log)",
                             "Population density (log)",
                             "Government expenditure (\\% GDP)",
                             "Trade (\\% of GDP)",
                             "Non agriculture value added (\\% GDP)",
                             "GDP growth",
                             "GDP per capita (log)",
                             "Population (log)"),
          single.row = FALSE, df = FALSE,
          omit = c("country", "year"),
          omit.stat = c("rsq", "f"),
          add.lines = list(c(rownames(summ.net.herf.2sls$diagnostics)[1], # use rownames as line name. in this case Weak insterumetns
                             round(summ.net.herf.2sls$diagnostics[1, "statistic"], 2),
                             round(summ.bb.herf.2sls$diagnostics[1, "statistic"], 2),
                             round(summ.fixed.herf.2sls$diagnostics[1, "statistic"], 2),
                             round(summ.mobile.herf.2sls$diagnostics[1, "statistic"], 2)),
                           c(rownames(summ.net.herf.2sls$diagnostics)[2],
                             round(summ.net.herf.2sls$diagnostics[2, "statistic"], 2),
                             round(summ.bb.herf.2sls$diagnostics[2, "statistic"], 2),
                             round(summ.fixed.herf.2sls$diagnostics[2, "statistic"], 2),
                             round(summ.mobile.herf.2sls$diagnostics[2, "statistic"], 2)),
                           c("P-value",
                             round(summ.net.herf.2sls$diagnostics[2, "p-value"], 2),
                             round(summ.bb.herf.2sls$diagnostics[2, "p-value"], 2),
                             round(summ.fixed.herf.2sls$diagnostics[2, "p-value"], 2),
                             round(summ.mobile.herf.2sls$diagnostics[2, "p-value"], 2)),
                           # c(rownames(summ.net.2sls$diagnostics)[3],
                           #   round(summ.net.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.bb.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.fixed.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.mobile.2sls$diagnostics[3, "statistic"], 2)),
                           # c("P-value",
                           #   round(summ.net.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.bb.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.fixed.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.mobile.2sls$diagnostics[3, "p-value"], 2)),
                           c("Country fixed effects", "Yes", "Yes", "Yes", "Yes"),
                           c("Yearly fixed effects", "Yes", "Yes", "Yes", "Yes")),
          notes = c("Robust Std. Errors in parenthesis",
                    "IV: Female participation in labour force"))

```

```{r, echo=FALSE, warning=FALSE, results='hide', message=FALSE}

# cv

net.cv.2sls <- ivreg(cv ~ net100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country) |
                      log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country)  +
                      fem.lab, 
               data = data, weights = 1/(se^2))
summary(net.cv.2sls, vcov = sandwich, diagnostics = TRUE)

bb.cv.2sls <- ivreg(cv ~ bb100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country) |
                      log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country)  +
                      fem.lab, 
               data = data, weights = 1/(se^2))
summary(bb.cv.2sls, vcov = sandwich, diagnostics = TRUE)

fixed.cv.2sls <- ivreg(cv ~ fixed100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country) |
                      log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country)  +
                      fem.lab, 
               data = data, weights = 1/(se^2))
summary(fixed.cv.2sls, vcov = sandwich, diagnostics = TRUE)

mobile.cv.2sls <- ivreg(cv ~ mobile100hab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country) |
                      log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country)  +
                      fem.lab, 
               data = data, weights = 1/(se^2))
summary(mobile.cv.2sls, vcov = sandwich, diagnostics = TRUE)

net.cv.2sls.rse <- coeftest(net.cv.2sls, function(x) vcovHC(x, type="HC0"))
summ.net.cv.2sls <- summary(net.cv.2sls, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)

bb.cv.2sls.rse <- coeftest(bb.cv.2sls, function(x) vcovHC(x, type="HC0"))
summ.bb.cv.2sls <- summary(bb.cv.2sls, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)

fixed.cv.2sls.rse <- coeftest(fixed.cv.2sls, function(x) vcovHC(x, type="HC0"))
summ.fixed.cv.2sls <- summary(fixed.cv.2sls, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)

mobile.cv.2sls.rse <- coeftest(mobile.cv.2sls, function(x) vcovHC(x, type="HC0"))
summ.mobile.cv.2sls <- summary(mobile.cv.2sls, vcov. = function(x) vcovHC(x, type="HC0"), diagnostics=T)
```

```{r, echo=FALSE, warning=FALSE, results='asis'}
stargazer(net.cv.2sls, bb.cv.2sls, fixed.cv.2sls, mobile.cv.2sls, type = "latex",
          se = list(net.cv.2sls.rse[,"Std. Error"], bb.cv.2sls.rse[,"Std. Error"],
                    fixed.cv.2sls.rse[,"Std. Error"], mobile.cv.2sls.rse[,"Std. Error"]),
          header = FALSE,
          no.space=TRUE,
          title = "2SLS estimation of equation (3), alt. specifications (iii)\\label{2sls.cv.global}",
          font.size = "small",
          column.sep.width = "1pt",
          dep.var.labels=c("Coefficient of variation 2000-18"),
          column.separate = c(1,1,1,1),
          covariate.labels=c("Internet users per 100 hab. (log)",
                             "Broadband users per 100 hab. (log)",
                             "Mobile phone users per 100 hab. (log)",
                             "Fixed phone users per 100 hab. (log)",
                             "Population density (log)",
                             "Government expenditure (\\% GDP)",
                             "Trade (\\% of GDP)",
                             "Non agriculture value added (\\% GDP)",
                             "GDP growth",
                             "GDP per capita (log)",
                             "Population (log)"),
          single.row = FALSE, df = FALSE,
          omit = c("country", "year"),
          omit.stat = c("rsq", "f"),
          add.lines = list(c(rownames(summ.net.cv.2sls$diagnostics)[1], # use rownames as line name. in this case Weak insterumetns
                             round(summ.net.cv.2sls$diagnostics[1, "statistic"], 2),
                             round(summ.bb.cv.2sls$diagnostics[1, "statistic"], 2),
                             round(summ.fixed.cv.2sls$diagnostics[1, "statistic"], 2),
                             round(summ.mobile.cv.2sls$diagnostics[1, "statistic"], 2)),
                           c(rownames(summ.net.cv.2sls$diagnostics)[2],
                             round(summ.net.cv.2sls$diagnostics[2, "statistic"], 2),
                             round(summ.bb.cv.2sls$diagnostics[2, "statistic"], 2),
                             round(summ.fixed.cv.2sls$diagnostics[2, "statistic"], 2),
                             round(summ.mobile.cv.2sls$diagnostics[2, "statistic"], 2)),
                           c("P-value",
                             round(summ.net.cv.2sls$diagnostics[2, "p-value"], 2),
                             round(summ.bb.cv.2sls$diagnostics[2, "p-value"], 2),
                             round(summ.fixed.cv.2sls$diagnostics[2, "p-value"], 2),
                             round(summ.mobile.cv.2sls$diagnostics[2, "p-value"], 2)),
                           # c(rownames(summ.net.2sls$diagnostics)[3],
                           #   round(summ.net.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.bb.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.fixed.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.mobile.2sls$diagnostics[3, "statistic"], 2)),
                           # c("P-value",
                           #   round(summ.net.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.bb.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.fixed.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.mobile.2sls$diagnostics[3, "p-value"], 2)),
                           c("Country fixed effects", "Yes", "Yes", "Yes", "Yes"),
                           c("Yearly fixed effects", "Yes", "Yes", "Yes", "Yes")),
          notes = c("Robust Std. Errors in parenthesis",
                    "IV: Female participation in labour force"))

```