### 2020-12-01 src/s2.Rmd

---
title: "S2 Appendix"
author: ""
output: pdf_document
header-includes:
- \usepackage{multirow}
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../paper/2_resubmission")
  })
---

```{r, echo=FALSE, warning=FALSE, results='hide', message=FALSE}
library(rmarkdown)
library(tidyverse)
library(car)
library(lmtest)
library(plm)
library(stargazer)
library(knitr)

un <- read_csv("C:/Users/nw19521/OneDrive - University of Bristol/projects/Ioannides/ict.un.us.uk/data/data_inter/data_for_un_regressions.csv")
us <- read_csv("C:/Users/nw19521/OneDrive - University of Bristol/projects/Ioannides/ict.un.us.uk/data/data_inter/data_for_usa_regressions.csv")
uk <- read_csv("C:/Users/nw19521/OneDrive - University of Bristol/projects/Ioannides/ict.un.us.uk/data/data_inter/data_for_uk_regressions.csv")
```

# First stage regressions for the 2SLS estimations

## Global model

```{r, echo=FALSE, warning=FALSE, results='asis', message=FALSE}

net.2sls.s1 <- lm(net100hab ~  fem.lab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country), 
               data = un, weights = 1/(se^2))

bb.2sls.s1 <- lm(bb100hab ~ fem.lab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country), 
               data = un, weights = 1/(se^2))

fixed.2sls.s1 <- lm(fixed100hab ~ fem.lab + log(pop.dens) + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country), 
               data = un, weights = 1/(se^2))

mobile.2sls.s1 <- lm(mobile100hab ~ fem.lab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country), 
               data = un, weights = 1/(se^2))

net.2sls.s1.rse <- coeftest(net.2sls.s1, function(x) vcovHC(x, type="HC0"))

bb.2sls.s1.rse <- coeftest(bb.2sls.s1, function(x) vcovHC(x, type="HC0"))

fixed.2sls.s1.rse <- coeftest(fixed.2sls.s1, function(x) vcovHC(x, type="HC0"))

mobile.2sls.s1.rse <- coeftest(mobile.2sls.s1, function(x) vcovHC(x, type="HC0"))

stargazer(net.2sls.s1, bb.2sls.s1, fixed.2sls.s1, mobile.2sls.s1, type = "latex",
          se = list(net.2sls.s1.rse[,"Std. Error"], bb.2sls.s1.rse[,"Std. Error"],
                    fixed.2sls.s1.rse[,"Std. Error"], mobile.2sls.s1.rse[,"Std. Error"]),
          header = FALSE,
          no.space=TRUE,
          title = "First stage regressions for Table 5 \\label{2sls.global.s1}",
          font.size = "footnotesize",
          column.sep.width = "1pt",
          # dep.var.labels=c("Internet users per 100 hab. (log)",
          #                 "Broadband users per 100 hab. (log)",
          #                 "Mobile phone users per 100 hab. (log)",
          #                 "Fixed phone users per 100 hab. (log)"),
          column.labels=c('Internet users',
                           "Broadband users",
                           "Mobile phone users",
                           "Fixed phone users"),
          dep.var.labels.include = FALSE,
          #column.separate = c(1,1,1,1),
          covariate.labels=c("Female labour force (\\%)",
                             "Population density (log)",
                             "Government expenditure (\\% GDP)",
                             "Trade (\\% of GDP)",
                             "Non agriculture value added (\\% GDP)",
                             "GDP growth",
                             "GDP per capita (log)",
                             "Population (log)"),
          single.row = FALSE, df = FALSE,
          omit = c("country", "year"),
          omit.stat = c("rsq", "f"),
          add.lines = list(
            # c(rownames(summ.net.2sls$diagnostics)[1], # use rownames as line name. in this case Weak insterumetns
            #                  round(summ.net.2sls$diagnostics[1, "statistic"], 2),
            #                  round(summ.bb.2sls$diagnostics[1, "statistic"], 2),
            #                  round(summ.fixed.2sls$diagnostics[1, "statistic"], 2),
            #                  round(summ.mobile.2sls$diagnostics[1, "statistic"], 2)),
            #                c(rownames(summ.net.2sls$diagnostics)[2],
            #                  round(summ.net.2sls$diagnostics[2, "statistic"], 2),
            #                  round(summ.bb.2sls$diagnostics[2, "statistic"], 2),
            #                  round(summ.fixed.2sls$diagnostics[2, "statistic"], 2),
            #                  round(summ.mobile.2sls$diagnostics[2, "statistic"], 2)),
            #                c("P-value",
            #                  round(summ.net.2sls$diagnostics[2, "p-value"], 2),
            #                  round(summ.bb.2sls$diagnostics[2, "p-value"], 2),
            #                  round(summ.fixed.2sls$diagnostics[2, "p-value"], 2),
            #                  round(summ.mobile.2sls$diagnostics[2, "p-value"], 2)),
                           # c(rownames(summ.net.2sls$diagnostics)[3],
                           #   round(summ.net.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.bb.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.fixed.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.mobile.2sls$diagnostics[3, "statistic"], 2)),
                           # c("P-value",
                           #   round(summ.net.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.bb.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.fixed.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.mobile.2sls$diagnostics[3, "p-value"], 2)),
                           c("Country fixed effects", "Yes", "Yes", "Yes", "Yes"),
                           c("Yearly fixed effects", "Yes", "Yes", "Yes", "Yes")),
          notes = c("Robust Std. Errors in parenthesis"))

net.test <- car::linearHypothesis(net.2sls.s1, c("fem.lab=0"))
bb.test <- car::linearHypothesis(bb.2sls.s1, c("fem.lab=0"))
mob.test <- car::linearHypothesis(fixed.2sls.s1, c("fem.lab=0"))
fixed.test <- car::linearHypothesis(mobile.2sls.s1, c("fem.lab=0"))

tests <- data.frame("F" = c(net.test$F[2], bb.test$F[2], mob.test$F[2], fixed.test$F[2]),
                    "p" = c(net.test$`Pr(>F)`[2], bb.test$`Pr(>F)`[2], mob.test$`Pr(>F)`[2], fixed.test$`Pr(>F)`[2]),
                    row.names = c("internet users", "broadband users", "mobile phone users", "fixed telephony users"))

stargazer(tests, 
          summary = F, 
          type = "latex",
          digits = 3,
          title = "F tests for IV (female labour force (\\%)) from first stage regressions for Table 5 \\label{2sls.global.s1.f}",
          header = FALSE
)
```

\newpage

## US model

```{r, echo=FALSE, warning=FALSE, results='asis', message=FALSE}

model_iv1.s1 <- lm(hh.net.share ~ bachelors2005 + 
                     log(pop2013) + un13 + white2013share + 
                     log(income2013) + I(pop2013/area_h) + service2013 + commuting.minutes2013, 
                   data = us)

model_iv2.s1 <- lm(hh.net.share ~ bachelors2005 + commuting.minutes2005 + 
                     log(pop2013) + un13 + white2013share + 
                     log(income2013) + I(pop2013/area_h) + service2013 + commuting.minutes2013, 
                   data = us)

rob.model_iv1.s1 <- coeftest(model_iv1.s1, function(x) vcovHC(x, type="HC0")) # rob = rse, naming inconsistency
rob.model_iv2.s1 <- coeftest(model_iv1.s1, function(x) vcovHC(x, type="HC0"))

stargazer(model_iv1.s1, model_iv2.s1, 
          type = "latex",
          title = "First stage regressions for Table 8 \\label{2sls.us.s1}",
          header = FALSE,
          se = list(rob.model_iv1.s1[,"Std. Error"], rob.model_iv2.s1[,"Std. Error"]), 
          dep.var.labels=c("\\% of households w. internet 2013"),
          covariate.labels=c("bachelors degree per hab. 2005",
                             "commute in minutes 2005",
                             "population 2013 (log)",
                             "\\% of unemployment 2013",
                             "\\% of white population 2013",
                             "income 2013 (log)",
                             "population density 2013",
                             "\\% of employment in service 2011",
                             "commute in minutes 2012"),
          single.row = FALSE,
          df = FALSE, 
          omit.stat = c("rsq", "f"), 
          notes = c("Robust Std. Error in parenthesis"))

model_iv1.s1.test <- car::linearHypothesis(model_iv1.s1, c("bachelors2005=0"))
model_iv2.s1.test <- car::linearHypothesis(model_iv2.s1, c("bachelors2005=0","commuting.minutes2005=0"))

tests <- data.frame("F" = c(model_iv1.s1.test$F[2], model_iv2.s1.test$F[2]),
                    "p" = c(model_iv1.s1.test$`Pr(>F)`[2], model_iv2.s1.test$`Pr(>F)`[2]),
                    row.names = c("bachelors degree per hab., 2005", "the above + commute in minutes, 2005"))

stargazer(tests, 
          summary = F, 
          type = "latex",
          digits = 3,
          title = "F tests for IVs from first stage regressions for Table 8 \\label{2sls.us.s1.f}",
          header = FALSE
)
```

\newpage

## UK model

```{r, echo=FALSE, warning=FALSE, results='asis', message=FALSE}

model_iv1.s1 <- lm(log(download2011) ~ uni.freq +
                     log(pop2011) + I(n.tests2011/pop2011) + unemployment.share + 
                     british.pop.share +  pop.dens + work.home.share + services.share
                     , data = uk, subset = n.tests2011>30)

model_iv2.s1 <- lm(log(download2011) ~ uni.freq + n.tests2011 +
                     log(pop2011) + I(n.tests2011/pop2011) + unemployment.share + 
                     british.pop.share +  pop.dens + work.home.share + services.share
                     , data = uk, subset = n.tests2011>30)

rob.model_iv1.s1 <- coeftest(model_iv1.s1, function(x) vcovHC(x, type="HC0"))
rob.model_iv2.s1 <- coeftest(model_iv2.s1, function(x) vcovHC(x, type="HC0"))

stargazer(model_iv1.s1, model_iv2.s1,
          type="latex",
          summary = F,
          header = FALSE,
          no.space=TRUE,
          title = "First stage regressions for Table 11 \\label{2sls.us.s1}",
          font.size = "small",
          se = list(rob.model_iv1.s1[,"Std. Error"], rob.model_iv2.s1[,"Std. Error"]),
          dep.var.labels=c("download speed, 2011 (log)"),
          covariate.labels=c("N. of universities",
                             "N. of broadband tests, 2011",
                             "population, 2011 (log)",  
                             "broadband tests per capita, 2011",
                             "\\% of unemployment, 2011",
                             "\\% of British population, 2011",
                             "population density, 2011",
                             "\\% of people working from home, 2011",
                             "employment in service, 2011 (\\%)"),
          single.row = FALSE,
          df = FALSE,
          omit.stat = c("rsq", "f"), 
          notes = c("Robust Std. Errors"))

# N. of universities",
# N. of universities, N. of broadband tests, 2011"),

model_iv1.s1.test <- car::linearHypothesis(model_iv1.s1, c("uni.freq=0"))
model_iv2.s1.test <- car::linearHypothesis(model_iv2.s1, c("uni.freq=0", "n.tests2011=0"))

tests <- data.frame("F" = c(model_iv1.s1.test$F[2], model_iv2.s1.test$F[2]),
                    "p" = c(model_iv1.s1.test$`Pr(>F)`[2], model_iv2.s1.test$`Pr(>F)`[2]),
                    row.names = c("N. of universities", "the above + N. of broadband tests, 2011"))

stargazer(tests, 
          summary = F, 
          type = "latex",
          digits = 3,
          title = "F tests for IVs from first stage regressions for Table 8 \\label{2sls.us.s1.f}",
          header = FALSE
)

```